---
title: "dbplot"
output:
  github_document:
    toc: true
    toc_depth: 3
---
```{r, setup, include = FALSE}
library(dplyr)
library(dbplot)
library(sparklyr)
library(nycflights13)

#Sys.setenv("JAVA_HOME" = normalizePath("C:\\Program Files (x86)\\Java\\jre1.8.0_144"))

knitr::opts_chunk$set(fig.height = 3.5, fig.width =  4, fig.align = 'center')
```


[![Build Status](https://travis-ci.org/edgararuiz/dbplot.svg?branch=master)](https://travis-ci.org/edgararuiz/dbplot)
[![CRAN\_Status\_Badge](http://www.r-pkg.org/badges/version/dbplot)](http://cran.r-project.org/package=dbplot)
![CRAN downloads](https://cranlogs.r-pkg.org/badges/grand-total/dbplot)

Leverages `dplyr` to process the calculations of a plot inside a database.  This package provides helper functions that abstract the work at three levels:
    
1. Functions that ouput a `ggplot2` object
2. Functions that outputs a `data.frame` object with the calculations
3. Creates the formula needed to calculate bins for a Histogram or a Raster plot

## Installation

```{r, eval = FALSE}
# You can install the released version from CRAN
install.packages("dbplot")

# Or the the development version from GitHub:
install.packages("devtools")
devtools::install_github("edgararuiz/dbplot")
```


## Connecting to a data source

- For more information on how to connect to databases, including Hive, please visit http://db.rstudio.com 

- To use Spark, please visit the `sparklyr` official website: http://spark.rstudio.com

## Example

In addition to database connections, the functions work with `sparklyr`. A Spark DataFrame will be used for the examples in this README.  

```{r, include = FALSE}
library(sparklyr)
sc <- spark_connect(master = "local", version = "2.1.0")
spark_flights <- copy_to(sc, nycflights13::flights, "flights")
```

```{r, eval = FALSE}
library(sparklyr)
sc <- spark_connect(master = "local", version = "2.1.0")
spark_flights <- copy_to(sc, nycflights13::flights, "flights")
```

## `ggplot`

### Histogram

By default `dbplot_histogram()` creates a 30 bin histogram

```{r}
library(ggplot2)

spark_flights %>% 
  dbplot_histogram(sched_dep_time)
```

Use `binwidth` to fix the bin size

```{r}
spark_flights %>% 
  dbplot_histogram(sched_dep_time, binwidth = 200)
```

Because it outputs a `ggplot2` object, more customization can be done

```{r}
spark_flights %>% 
  dbplot_histogram(sched_dep_time, binwidth = 300) +
  labs(title = "Flights - Scheduled Departure Time") +
  theme_bw()
```

### Raster

To visualize two continuous variables, we typically resort to a Scatter plot. However, this may not be practical when visualizing millions or billions of dots representing the intersections of the two variables. A Raster plot may be a better option, because it concentrates the intersections into squares that are easier to parse visually.

A Raster plot basically does the same as a Histogram. It takes two continuous variables and creates discrete 2-dimensional bins represented as squares in the plot. It then determines either the number of rows inside each square or processes some aggregation, like an average.


- If no `fill` argument is passed, the default calculation will be count, `n()`
```{r}
spark_flights %>%
  filter(!is.na(arr_delay)) %>%
  dbplot_raster(arr_delay, dep_delay) 
```


- Pass an aggregation formula that can run inside the database
```{r}
spark_flights %>%
  filter(!is.na(arr_delay)) %>%
  dbplot_raster(arr_delay, dep_delay, mean(distance, na.rm = TRUE)) 
```

- Increase or decrease for more, or less, definition.  The `resolution` argument controls that, it defaults to 100 
```{r}
spark_flights %>%
  filter(!is.na(arr_delay)) %>%
  dbplot_raster(arr_delay, dep_delay, mean(distance, na.rm = TRUE), resolution = 500)
```

### Bar Plot

- `dbplot_bar()` defaults to a tally() of each value in a discrete variable
```{r}
spark_flights %>%
  dbplot_bar(origin)
```


- Pass a formula that will be operated for each value in the discrete variable
```{r}
spark_flights %>%
  dbplot_bar(origin, mean(dep_delay))
```

### Line plot

- `dbplot_line()` defaults to a tally() of each value in a discrete variable
```{r}
spark_flights %>%
  dbplot_line(month)
```

- Pass a formula that will be operated for each value in the discrete variable
```{r}
spark_flights %>%
  dbplot_line(month, mean(dep_delay))
```

### Boxplot

- It expects a discrete variable to group by, and a continuous variable to calculate the percentiles and IQR. It doesn't calculate outliers. Currently, this feature works with sparklyr and Hive connections.

```{r}
spark_flights %>%
  dbplot_boxplot(origin, dep_delay)
```


## Calculation functions

If a more customized plot is needed, the data the underpins the plots can also be accessed:

1. `db_compute_bins()` - Returns a data frame with the bins and count per bin
2. `db_compute_count()` - Returns a data frame with the count per discrete value
3. `db_compute_raster()` -  Returns a data frame with the results per x/y intersection
4. `db_compute_boxplot()` -  Returns a data frame with boxplot calculations


```{r}
spark_flights %>%
  db_compute_bins(arr_delay) 
```

The data can be piped to a plot
```{r}
spark_flights %>%
  filter(arr_delay < 100 , arr_delay > -50) %>%
  db_compute_bins(arr_delay) %>%
  ggplot() +
  geom_col(aes(arr_delay, count, fill = count))
```


## `db_bin()`

Uses 'rlang' to build the formula needed to create the bins of a numeric variable in an un-evaluated fashion. This way, the formula can be then passed inside a dplyr verb.

```{r}
db_bin(var)
```


```{r}
spark_flights %>%
  group_by(x = !! db_bin(arr_delay)) %>%
  tally()
```

```{r}
spark_flights %>%
  filter(!is.na(arr_delay)) %>%
  group_by(x = !! db_bin(arr_delay)) %>%
  tally()%>%
  collect %>%
  ggplot() +
  geom_col(aes(x, n))
```


